---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  # out.width = "100%",
  fig.retina = 2
)
```

# ggdirectlabel

<!-- badges: start -->
[![R-CMD-check](https://github.com/MattCowgill/ggdirectlabel/workflows/R-CMD-check/badge.svg)](https://github.com/MattCowgill/ggdirectlabel/actions)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/MattCowgill/ggdirectlabel/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/MattCowgill/ggdirectlabel/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of ggdirectlabel is to make it easier to directly label ggplot2 charts rather than using legends.

## Installation

You can install the development version of ggdirectlabel from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("MattCowgill/ggdirectlabel")
```

```{r}
library(ggdirectlabel)
library(ggplot2)
library(magrittr)
```

## Using `geom_richlegend()`
Here's a standard ggplot2 scatterplot:
```{r}
base_scatter <- mtcars |> 
  ggplot(aes(x = wt, y = mpg, col = factor(cyl))) +
  geom_point()

base_scatter
```

This is fine! But sometimes you might like the legend levels (4, 6, and 8 in this example) to be coloured according to the levels in the data. That's where `geom_richlegend()` comes in:
```{r}
base_scatter +
  geom_richlegend(aes(label = cyl)) +
  theme(legend.position = "none")
```

You can move the 'rich legend' around:
```{r}
base_scatter +
  geom_richlegend(aes(label = cyl),
                  legend.position = "bottomleft",
                  vjust = 0,
                  hjust = 0) +
  theme(legend.position = "none")
```

`geom_richlegend()` respects facets - it'll place a little legend annotation for each level of the data that appears in that panel:

```{r}
base_scatter +
  geom_richlegend(aes(label = paste0(cyl, " cylinders"))) +
  facet_wrap(~cyl)
```


## Using `geom_linepoint()`

Without ggirectlabel, we might do something like:

```{r no-directlabel}

ggplot2::economics_long %>%
  ggplot(aes(x = date, y = value, col = variable)) +
  geom_line() +
  geom_point(data = ~dplyr::filter(., date == max(date)),
             fill = "white",
             shape = 21,
             size = 2.5,
             stroke = 1.25)
```

This is fine! But this is a more straightforward way to achieve the same thing:

```{r example}
ggplot2::economics_long %>%
  ggplot(aes(x = date, y = value, col = variable)) +
  geom_linepoint()
```

## Using `scale_x_date_rightalign()`

In time series line charts, it's often important to make clear the date of 
your most recent observation. The `scale_x_date_rightalign()` function aligns
the breaks of your x-axis so that the most recent observation is included in 
the breaks.

```{r scale_x_date_rightalign}
ggplot2::economics_long %>%
  ggplot(aes(x = date, y = value, col = variable)) +
  geom_linepoint() +
  scale_x_date_rightalign()
```

## Using `geom_finallabel()`

In time series line charts, you may wish to label the final point in the series. The `geom_finallabel()` function makes that easy. 

```{r geom_finallabel}
ggplot2::economics_long %>%
  ggplot(aes(x = date, y = value, col = variable)) +
  geom_linepoint() +
  geom_finallabel(aes(label = round(value, 0))) +
  scale_x_date_rightalign(expand = expansion(c(0, 0.15))) +
  theme(legend.position = "none")
```

